---
title: "Get stats for days relative to CRS"
author: Fiorella Schischlik
date: June, 30 2020
output:
  md_document:
    variant: gfm
---

load libraries.
```{r}
library(tidyverse)
```

Directories.
```{r}
# dir_data <- "~/datasets/CAR-T/figs_main/"
# for revision
dir_data <- "~/src/2020-CAR22-toxicities/figures/figs_revision/"
other_cytokines <- "~/src/2020-CAR22-toxicities/data/IL-18BP/"
dir_crs <- "~/src/2020-CAR22-toxicities/data/"
```

Load and merge datasets together.
```{r}

# New ASBMT CRS grading
asbmt_grading <- read_csv(paste(dir_crs, "ID Key_15C0029 CD22 CAR CRS Data_4-30-21_export.csv", sep=""))

crs <- read_csv(paste(dir_data, "day_relative_to_CRS_export.csv", sep="")) %>%
  dplyr::select(-X1)

il18bp <- readxl::read_xlsx(paste(other_cytokines, "CD22_IL18bp_12-27-20_v1.xlsx", sep=""))


il18bp <- il18bp %>% dplyr::rename(
  patient_id = Subject_ID,
  max_grade_CRS = `Max Grade CRS`,
  date = `Day Post Infusion`,
  days_rel_to_CRS = `Day Relative to CRS Onset`,
  cytokine_levels = `IL-18BP pg/nl`) %>%
  dplyr::mutate(
    index=NA,
    date_CRS=NA,
    CRS=CRS,
    cytokines="IL-18BP",
    days_HLH=NA,
    log10_cytokine_levels=log10(cytokine_levels),
    `CRS grade` = ifelse(`max_grade_CRS` %in% c("1", "2"), "CRS 1-2", "CRS 3-4")) %>%  
  dplyr::select(
    colnames(crs)) %>%
  dplyr::filter(days_rel_to_CRS %in% as.character(unique(crs$days_rel_to_CRS)))

# Recode HLH variable
il18bp$HLH <- dplyr::recode_factor(il18bp$HLH, `1`="carHLH+", `0`="carHLH-")

cytokine_table <- rbind(il18bp, crs)

```

Mail from Nirali:
 - Calculate a separate table: with median, mean, IQR and SD with stats 
 for each timepoint to determine difference, and the number evaluable at 
 each timepoint (stratified by HLH and not-HLH)
- Are you able to do this table in BOTH: Actual value and log10 of cytokines
```{r}

# test
timepoint <- 0
cytokine <- "IL-18BP"

# Significant timepoints
timepoints <- sort(unique(cytokine_table$days_rel_to_CRS))
cytokines <- unique(cytokine_table$cytokines)

# initialize
summary_table <- data.frame(
  days_rel_to_CRS = integer(),
  cytokine = character(),
  num_patients = numeric(),
  carHLH_pos = numeric(),
  carHLH_neg = numeric(),
  carHLH_pos_mean = numeric(),
  carHLH_neg_mean = numeric(),
  carHLH_pos_log10_mean = numeric(),
  carHLH_neg_log10_mean = numeric(),
  carHLH_pos_log10_median = numeric(),
  carHLH_neg_log10_median = numeric(),
  carHLH_pos_log10_IQR = numeric(),
  carHLH_neg_log10_IQR = numeric(),
  wilcox_pvalue = numeric()
)

summary_table$cytokine <- as.character(summary_table$cytokine)

i <- 1

for (timepoint in timepoints){
  print(timepoint)
  for (cytokine in cytokines){
    print(cytokine)

    df <- cytokine_table %>% dplyr::filter(days_rel_to_CRS==timepoint & cytokines==cytokine)
    non_na <- df[complete.cases(df$cytokine_levels), ]
    
    if(dim(non_na)[1]!=0 & (! 0 %in% table(non_na$HLH))){
      
      carHLH_pos_cytokine_levels <- dplyr::filter(non_na, HLH=="carHLH+" )$cytokine_levels
      carHLH_neg_cytokine_levels <- dplyr::filter(non_na, HLH=="carHLH-" )$cytokine_levels
  
      wilcox_res <- wilcox.test(
        carHLH_pos_cytokine_levels,
        carHLH_neg_cytokine_levels,
        alternative = c("two.sided"))
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$num_patients <- sum(length(carHLH_neg_cytokine_levels) + length(carHLH_pos_cytokine_levels))
      summary_table[i,]$carHLH_pos <- length(carHLH_pos_cytokine_levels)
      summary_table[i,]$carHLH_neg <- length(carHLH_neg_cytokine_levels)
      summary_table[i,]$carHLH_pos_mean <- mean(carHLH_pos_cytokine_levels)
      summary_table[i,]$carHLH_neg_mean <- mean(carHLH_neg_cytokine_levels)
      summary_table[i,]$carHLH_pos_log10_mean <- mean(log10(carHLH_pos_cytokine_levels))
      summary_table[i,]$carHLH_neg_log10_mean <- mean(log10(carHLH_neg_cytokine_levels))
      summary_table[i,]$carHLH_pos_log10_median <- median(log10(carHLH_pos_cytokine_levels))
      summary_table[i,]$carHLH_neg_log10_median <- median(log10(carHLH_neg_cytokine_levels))
      summary_table[i,]$carHLH_pos_log10_IQR <- IQR(log10(carHLH_pos_cytokine_levels))
      summary_table[i,]$carHLH_neg_log10_IQR <- IQR(log10(carHLH_neg_cytokine_levels))
      summary_table[i,]$wilcox_pvalue <- wilcox_res$p.value
        
    }else{
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$num_patients <- NA
      summary_table[i,]$carHLH_pos <- NA
      summary_table[i,]$carHLH_neg <- NA
      summary_table[i,]$carHLH_pos_mean <- NA
      summary_table[i,]$carHLH_neg_mean <- NA
      summary_table[i,]$carHLH_pos_log10_mean <- NA
      summary_table[i,]$carHLH_neg_log10_mean <- NA
      summary_table[i,]$carHLH_pos_log10_median <- NA
      summary_table[i,]$carHLH_neg_log10_median <- NA
      summary_table[i,]$carHLH_pos_log10_IQR <- NA
      summary_table[i,]$carHLH_neg_log10_IQR <- NA
      summary_table[i,]$wilcox_pvalue <- NA
    }
    
    i <- i + 1
    
    #print(df)
  }
}


# Get fold change
summary_table$log2fc <- log2(summary_table$carHLH_pos_mean/summary_table$carHLH_neg_mean)

# Adjust using FDR 
summary_table$padj_FDR <- NA 

for (timepoint in timepoints){
  df_timepoint <- summary_table %>% dplyr::filter(days_rel_to_CRS==timepoint)
  summary_table[summary_table$days_rel_to_CRS==timepoint, ]$padj_FDR <- p.adjust(df_timepoint$wilcox_pvalue, method="BH")
}

write_excel_csv(
  summary_table, 
  file=paste(dir_data, "rel_to_CRS_stats_carHLH_with_pvalues.csv", sep=""))

```

Calculate a separate table: with median, mean, IQR and SD with stats 
 for each timepoint to determine difference, and the number evaluable at 
 each timepoint (stratified by CRS 1-2 and CRS 3-4)
```{r}

cytokine_table$log10_cytokine_levels <- log10(cytokine_table$cytokine_levels)
  
# Significant timepoints
timepoints <- sort(unique(cytokine_table$days_rel_to_CRS))
cytokines <- unique(cytokine_table$cytokines)

# initialize
summary_table <- data.frame(
  days_rel_to_CRS = integer(),
  cytokine = character(),
  num_patients = numeric(),
  CRS_low = numeric(),
  CRS_high = numeric(),
  CRS_low_mean = numeric(),
  CRS_high_mean = numeric(),
  CRS_low_log10_mean = numeric(),
  CRS_high_log10_mean = numeric(),
  CRS_low_log10_median = numeric(),
  CRS_high_log10_median = numeric(),
  CRS_low_log10_IQR = numeric(),
  CRS_high_log10_IQR = numeric(),
  wilcox_pvalue = numeric()
)

summary_table$cytokine <- as.character(summary_table$cytokine)

i <- 1

for (timepoint in timepoints){
  print(timepoint)
  for (cytokine in cytokines){
    print(cytokine)

    df <- cytokine_table %>% dplyr::filter(days_rel_to_CRS==timepoint & cytokines==cytokine)
    non_na <- df[complete.cases(df$cytokine_levels), ]
    
    if(dim(non_na)[1]!=0 & (length(table(non_na$`CRS grade`))==2)){
  
      CRS_low_cytokine_levels <- dplyr::filter(non_na, `CRS grade`=="CRS 1-2" )$cytokine_levels
      CRS_high_cytokine_levels <- dplyr::filter(non_na, `CRS grade`=="CRS 3-4" )$cytokine_levels
      
      wilcox_res <- wilcox.test(
        CRS_low_cytokine_levels,
        CRS_high_cytokine_levels,
        alternative = c("two.sided"))
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$wilcox_pvalue <- wilcox_res$p.value
      summary_table[i,]$num_patients <- sum(table(non_na$`CRS grade`)[1] + table(non_na$`CRS grade`)[2])
      summary_table[i,]$CRS_low <- table(non_na$`CRS grade`)[1]
      summary_table[i,]$CRS_high <- table(non_na$`CRS grade`)[2]
      summary_table[i,]$CRS_low_mean <- mean(CRS_low_cytokine_levels)
      summary_table[i,]$CRS_high_mean <- mean(CRS_high_cytokine_levels)
      summary_table[i,]$CRS_low_log10_mean <- mean(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_mean <- mean(log10(CRS_high_cytokine_levels))
      summary_table[i,]$CRS_low_log10_median <- median(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_median <- median(log10(CRS_high_cytokine_levels))
      summary_table[i,]$CRS_low_log10_IQR <- IQR(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_IQR <- IQR(log10(CRS_high_cytokine_levels))
        
    }else{
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$wilcox_pvalue <- NA
      summary_table[i,]$num_patients <- NA
      summary_table[i,]$CRS_low <- NA
      summary_table[i,]$CRS_high <- NA
      summary_table[i,]$CRS_low_log10_mean <- NA
      summary_table[i,]$CRS_high_log10_mean <- NA
      summary_table[i,]$CRS_low_log10_median <- NA
      summary_table[i,]$CRS_high_log10_median <- NA
      summary_table[i,]$CRS_low_log10_IQR <- NA
      summary_table[i,]$CRS_high_log10_IQR <- NA
    }
    
    i <- i + 1
    
    #print(df)
  }
}


# Get fold change
summary_table$log2fc <- log2(summary_table$CRS_high_mean/summary_table$CRS_low_mean)

# Adjust using FDR 
summary_table$padj_FDR <- NA 

for (timepoint in timepoints){
  df_timepoint <- summary_table %>% dplyr::filter(days_rel_to_CRS==timepoint)
  summary_table[summary_table$days_rel_to_CRS==timepoint, ]$padj_FDR <- p.adjust(df_timepoint$wilcox_pvalue, method="BH")
}


write_excel_csv(
  summary_table, 
  path=paste(dir_data, "rel_to_CRS_stats_CRS_grade_with_pvalues.csv", sep=""))

```

Merge patient with new CRS grading system (ASBMT).
```{r}

cytokine_table_merged <- merge(
  cytokine_table, 
  asbmt_grading, 
  by.x = "patient_id",
  by.y = "Randomized ID")

cytokine_table_merged <- as_tibble(
  cytokine_table_merged) %>%
  dplyr::mutate(crs_grade_asbmt = ifelse(`Max Grade CRS_ASBMT` %in% c("1", "2"), 
                                        "CRS 1-2", "CRS 3-4"))

```

Calculate a separate table: with median, mean, IQR and SD with stats 
 for each timepoint to determine difference, and the number evaluable at 
 each timepoint (stratified by CRS 1-2 and CRS 3-4) use CRS ASBMT grading
```{r}

cytokine_table <- cytokine_table_merged
cytokine_table$log10_cytokine_levels <- log10(cytokine_table$cytokine_levels)
  
# Significant timepoints
timepoints <- sort(unique(cytokine_table$days_rel_to_CRS))
cytokines <- unique(cytokine_table$cytokines)

# initialize
summary_table <- data.frame(
  days_rel_to_CRS = integer(),
  cytokine = character(),
  num_patients = numeric(),
  CRS_low = numeric(),
  CRS_high = numeric(),
  CRS_low_mean = numeric(),
  CRS_high_mean = numeric(),
  CRS_low_log10_mean = numeric(),
  CRS_high_log10_mean = numeric(),
  CRS_low_log10_median = numeric(),
  CRS_high_log10_median = numeric(),
  CRS_low_log10_IQR = numeric(),
  CRS_high_log10_IQR = numeric(),
  wilcox_pvalue = numeric()
)

summary_table$cytokine <- as.character(summary_table$cytokine)

i <- 1

for (timepoint in timepoints){
  print(timepoint)
  for (cytokine in cytokines){
    print(cytokine)

    df <- cytokine_table %>% dplyr::filter(days_rel_to_CRS==timepoint & cytokines==cytokine)
    non_na <- df[complete.cases(df$cytokine_levels), ]
    
    if(dim(non_na)[1]!=0 & (length(table(non_na$`crs_grade_asbmt`))==2)){
  
      CRS_low_cytokine_levels <- dplyr::filter(non_na, `crs_grade_asbmt`=="CRS 1-2" )$cytokine_levels
      CRS_high_cytokine_levels <- dplyr::filter(non_na, `crs_grade_asbmt`=="CRS 3-4" )$cytokine_levels
      
      wilcox_res <- wilcox.test(
        CRS_low_cytokine_levels,
        CRS_high_cytokine_levels,
        alternative = c("two.sided"))
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$wilcox_pvalue <- wilcox_res$p.value
      summary_table[i,]$num_patients <- sum(table(non_na$`crs_grade_asbmt`)[1] + table(non_na$`crs_grade_asbmt`)[2])
      summary_table[i,]$CRS_low <- table(non_na$`crs_grade_asbmt`)[1]
      summary_table[i,]$CRS_high <- table(non_na$`crs_grade_asbmt`)[2]
      summary_table[i,]$CRS_low_mean <- mean(CRS_low_cytokine_levels)
      summary_table[i,]$CRS_high_mean <- mean(CRS_high_cytokine_levels)
      summary_table[i,]$CRS_low_log10_mean <- mean(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_mean <- mean(log10(CRS_high_cytokine_levels))
      summary_table[i,]$CRS_low_log10_median <- median(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_median <- median(log10(CRS_high_cytokine_levels))
      summary_table[i,]$CRS_low_log10_IQR <- IQR(log10(CRS_low_cytokine_levels))
      summary_table[i,]$CRS_high_log10_IQR <- IQR(log10(CRS_high_cytokine_levels))
        
    }else{
      
      summary_table[i,]$days_rel_to_CRS <- timepoint
      summary_table[i,]$cytokine <- cytokine
      summary_table[i,]$wilcox_pvalue <- NA
      summary_table[i,]$num_patients <- NA
      summary_table[i,]$CRS_low <- NA
      summary_table[i,]$CRS_high <- NA
      summary_table[i,]$CRS_low_log10_mean <- NA
      summary_table[i,]$CRS_high_log10_mean <- NA
      summary_table[i,]$CRS_low_log10_median <- NA
      summary_table[i,]$CRS_high_log10_median <- NA
      summary_table[i,]$CRS_low_log10_IQR <- NA
      summary_table[i,]$CRS_high_log10_IQR <- NA
    }
    
    i <- i + 1
    
    #print(df)
  }
}


# Get fold change
summary_table$log2fc <- log2(summary_table$CRS_high_mean/summary_table$CRS_low_mean)

# Adjust using FDR 
summary_table$padj_FDR <- NA 

for (timepoint in timepoints){
  df_timepoint <- summary_table %>% dplyr::filter(days_rel_to_CRS==timepoint)
  summary_table[summary_table$days_rel_to_CRS==timepoint, ]$padj_FDR <- p.adjust(df_timepoint$wilcox_pvalue, method="BH")
}


write_excel_csv(
  summary_table, 
  path=paste(dir_data, "rel_to_CRS_stats_CRS_ASBMT_grade_with_pvalues.csv", sep=""))

```
